---
title: "Introduction to gkwreg2"
author: "José Evandeilton Lopes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gkwreg2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 7,
    fig.height = 5
)
```

## Overview

The **gkwreg2** package provides regression modeling for bounded data in (0, 1) 
using the Generalized Kumaraswamy distribution family. This includes proportions, 
rates, percentages, and any continuous data restricted to the unit interval.

## Installation

```{r eval=FALSE}
# Install from GitHub
remotes::install_github("evandeilton/gkwdist")
remotes::install_github("evandeilton/gkwreg2")
```

## Distribution Families

The package supports 7 distribution families:
- **gkw**: Generalized Kumaraswamy (5 parameters)
- **bkw**: Beta-Kumaraswamy (4 parameters)
- **kkw**: Kumaraswamy-Kumaraswamy (4 parameters)
- **ekw**: Exponentiated Kumaraswamy (3 parameters)
- **mc**: McDonald/Beta Power (3 parameters)
- **kw**: Kumaraswamy (2 parameters)
- **beta**: Beta (2 parameters)

## Quick Example

```{r example, message=FALSE, warning=FALSE}
library(gkwreg2)
library(gkwdist)

# Simulate data
set.seed(123)
n <- 300
x1 <- runif(n, -2, 2)
x2 <- rnorm(n)

# Generate response with known parameters
alpha_true <- exp(0.8 + 0.3 * x1)
beta_true <- exp(1.2 - 0.2 * x2)
y <- rkw(n, alpha = alpha_true, beta = beta_true)
y <- pmax(pmin(y, 1 - 1e-7), 1e-7)

df <- data.frame(y = y, x1 = x1, x2 = x2)

# Fit Kumaraswamy regression
# alpha depends on x1, beta depends on x2
fit <- gkwreg2(y ~ x1 | x2, data = df, family = "kw")

summary(fit)
```

## Extended Formula Syntax

The pipe `|` separator allows specifying different covariates for each parameter:

```{r formula}
# Intercept-only model
fit0 <- gkwreg2(y ~ 1, data = df, family = "kw")

# x1 for alpha only (beta = intercept)
fit1 <- gkwreg2(y ~ x1, data = df, family = "kw")

# x1 for alpha, x2 for beta
fit2 <- gkwreg2(y ~ x1 | x2, data = df, family = "kw")

# Same covariates for both
fit3 <- gkwreg2(y ~ x1 + x2 | x1 + x2, data = df, family = "kw")
```

## Model Comparison

```{r comparison}
# Compare models using AIC
AIC(fit0, fit1, fit2, fit3)

# Likelihood ratio tests
anova(fit0, fit1, fit2, test = "Chisq")
```

## Predictions

```{r predictions}
# Predicted means
pred_mean <- predict(fit2, type = "response")

# Distribution parameters
pred_params <- predict(fit2, type = "parameters")
head(data.frame(alpha = pred_params$alpha, beta = pred_params$beta))

# Quantiles
pred_q <- predict(fit2, type = "quantile", at = c(0.1, 0.5, 0.9))
head(pred_q)
```

## Diagnostic Plots

```{r diagnostics, fig.height=8}
par(mfrow = c(2, 2))
plot(fit2, which = c(2, 3, 5, 6), ask = FALSE)
```

## Comparing Distribution Families

```{r families, warning=FALSE}
# Fit different families
fit_kw <- gkwreg2(y ~ x1 | x2, data = df, family = "kw")
fit_beta <- gkwreg2(y ~ x1 | x2, data = df, family = "beta")

# Compare
data.frame(
    Family = c("Kumaraswamy", "Beta"),
    LogLik = c(logLik(fit_kw), logLik(fit_beta)),
    AIC = c(AIC(fit_kw), AIC(fit_beta)),
    BIC = c(BIC(fit_kw), BIC(fit_beta))
)
```

## Optimization Control

Fine-tune the optimization with `gkw_control()`:

```{r control, eval=FALSE}
# Use BFGS optimizer with more iterations
fit <- gkwreg2(y ~ x1,
    data = df, family = "kw",
    control = gkw_control(method = "BFGS", maxit = 500)
)

# Fast fitting without standard errors
fit_fast <- gkwreg2(y ~ x1,
    data = df, family = "kw",
    control = gkw_control(hessian = FALSE)
)
```

## See Also

- `?gkwreg2` - Main fitting function
- `?gkw_control` - Optimization control
- `?predict.gkwreg2` - Prediction types
- `?plot.gkwreg2` - Diagnostic plots
